---
title: "Sample Characteristics"
---



```{r}
library(dplyr) # Run install.packages("tidyverse") to get this
library(tidyr) # Run install.packages("tidyr") to get this
library(ggplot2)
library("reshape") # Run install.packages("reshape") to get this
```

Import the csv with characteristics and other info from GEO sample page

```{r}

sample_characteristics <- read.csv("alcohol_hippocampus_samples.csv", header = TRUE)

```

Save n for calculations

```{r}

n_samples <- nrow(sample_characteristics)

```


How many alcoholics to controls?

```{r}

n_alcoholics <- sum(sample_characteristics$phenotype == "alcoholic")
n_controls <- sum(sample_characteristics$phenotype == "control")

print(paste("% alcoholic = ", (n_alcoholics / n_samples)*100))
print(paste("% controls = ", (n_controls / n_samples)*100))

```
Look at sex

```{r}

n_female <- sum(sample_characteristics$sex == "female")
n_male <- sum(sample_characteristics$sex == "male")

print(paste("% female = ", (n_female / n_samples)*100))
print(paste("% male = ", (n_male / n_samples)*100))

```
```{r}

female_alcoholic <- sample_characteristics %>%
  filter(sex == "female") %>%
  filter(phenotype == "alcoholic")

male_alcoholic <- sample_characteristics %>%
  filter(sex == "male") %>%
  filter(phenotype == "alcoholic")


print(paste("% alcoholics in female = ", (nrow(female_alcoholic) / n_female)*100))
print(paste("% alcoholics in male = ", (nrow(male_alcoholic) / n_male)*100))

```
Look at age

```{r}

print(paste("Min age:", min(sample_characteristics$age)))
print(paste("Max age:", max(sample_characteristics$age)))
print(paste("Median age:", median(sample_characteristics$age)))

```

Check the percentages with cirrhosis

```{r}

print(paste("N cirrhosis:", sum(sample_characteristics$cirrhosis == "yes")))
print(paste("% cirrhosis = ", (sum(sample_characteristics$cirrhosis == "yes") / n_samples)*100))

```
Check smoking

```{r}

print(paste("N active smoker:", sum(sample_characteristics$smoker == "yes")))
print(paste("N former smoker:", sum(sample_characteristics$smoker == "former")))
print(paste("N non smoker:", sum(sample_characteristics$smoker == "no")))
print(paste("N unknown smoker:", sum(sample_characteristics$smoker == "unknown")))


print(paste("% active smoker = ", (sum(sample_characteristics$smoker == "yes") / n_samples)*100))
print(paste("% former smoker = ", (sum(sample_characteristics$smoker == "former") / n_samples)*100))
print(paste("% non smoker = ", (sum(sample_characteristics$smoker == "no") / n_samples)*100))
print(paste("% unknown = ", (sum(sample_characteristics$smoker == "unknown") / n_samples)*100))

```

Look at post mortem interval

```{r}

print(paste("Min interval:", min(sample_characteristics$PMI)))
print(paste("Max interval:", max(sample_characteristics$PMI)))
print(paste("Median interval:", median(sample_characteristics$PMI)))


```
```{r}

print(paste("Median brain pH:", median(sample_characteristics$brain.pH)))

```

Get the values again with respect to phenotype so that we can know if there are significant differences between them other than alcohol use.

```{r}

alcoholic <- sample_characteristics %>%
  filter(phenotype == "alcoholic")

control <- sample_characteristics %>%
  filter(phenotype == "control")

```


=============================================================
ALcoholic

```{r}

n_female <- sum(alcoholic$sex == "female")
n_male <- sum(alcoholic$sex == "male")

print(paste("% female = ", (n_female / nrow(alcoholic))*100))
print(paste("% male = ", (n_male / nrow(alcoholic))*100))

```

Look at age

```{r}

print(paste("Min age:", min(alcoholic$age)))
print(paste("Max age:", max(alcoholic$age)))
print(paste("Median age:", median(alcoholic$age)))

```

Check the percentages with cirrhosis

```{r}

print(paste("N cirrhosis:", sum(alcoholic$cirrhosis == "yes")))
print(paste("% cirrhosis = ", (sum(alcoholic$cirrhosis == "yes") / nrow(alcoholic))*100))

```
Check smoking

```{r}

print(paste("N active smoker:", sum(alcoholic$smoker == "yes")))
print(paste("N former smoker:", sum(alcoholic$smoker == "former")))
print(paste("N non smoker:", sum(alcoholic$smoker == "no")))
print(paste("N unknown smoker:", sum(alcoholic$smoker == "unknown")))


print(paste("% active smoker = ", (sum(alcoholic$smoker == "yes") / n_samples)*100))
print(paste("% former smoker = ", (sum(alcoholic$smoker == "former") / n_samples)*100))
print(paste("% non smoker = ", (sum(alcoholic$smoker == "no") / n_samples)*100))
print(paste("% unknown = ", (sum(alcoholic$smoker == "unknown") / n_samples)*100))

```

Look at post mortem interval

```{r}

print(paste("Min interval:", min(alcoholic$PMI)))
print(paste("Max interval:", max(alcoholic$PMI)))
print(paste("Median interval:", median(alcoholic$PMI)))


```


```{r}

print(paste("Median brain pH:", median(alcoholic$brain.pH)))

```



====================================================================
Controls

```{r}

n_female <- sum(control$sex == "female")
n_male <- sum(control$sex == "male")

print(paste("% female = ", (n_female / nrow(control))*100))
print(paste("% male = ", (n_male / nrow(control))*100))

```

Look at age

```{r}

print(paste("Min age:", min(control$age)))
print(paste("Max age:", max(control$age)))
print(paste("Median age:", median(control$age)))

```

Check the percentages with cirrhosis

```{r}

print(paste("N cirrhosis:", sum(control$cirrhosis == "yes")))
print(paste("% cirrhosis = ", (sum(control$cirrhosis == "yes") / nrow(control))*100))

```
Check smoking

```{r}

print(paste("N active smoker:", sum(control$smoker == "yes")))
print(paste("N former smoker:", sum(control$smoker == "former")))
print(paste("N non smoker:", sum(control$smoker == "no")))
print(paste("N unknown smoker:", sum(control$smoker == "unknown")))


print(paste("% active smoker = ", (sum(control$smoker == "yes") / n_samples)*100))
print(paste("% former smoker = ", (sum(control$smoker == "former") / n_samples)*100))
print(paste("% non smoker = ", (sum(control$smoker == "no") / n_samples)*100))
print(paste("% unknown = ", (sum(control$smoker == "unknown") / n_samples)*100))

```

Look at post mortem interval

```{r}

print(paste("Min interval:", min(control$PMI)))
print(paste("Max interval:", max(control$PMI)))
print(paste("Median interval:", median(control$PMI)))


```


```{r}

print(paste("Median brain pH:", median(control$brain.pH)))

```








